---
date: 
external_link: 
image:
  caption: Cases over in last 2 Weeks
  focal_point: Smart
summary: Cases by County last 2 Weeks.
tags:
- Demo
title: Cases by County last 2 Weeks.
---
<br><br><br><br>

Breaking down the data to the level of cases per county over the last 2 weeks allows us to observe the current infection hotspots nationwide in a manner which is obscured when viewing the total amount of cases per county.

<br><br><br><br>

```{r setup, include=FALSE}

rm(list=ls())
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(RCurl)
x <- getURL("https://raw.githubusercontent.com/frankmoriarty/covid_ie/master/doh_covid_ie_cases_analysis.csv")
newcases <- read.csv(text = x, stringsAsFactors=FALSE)

h <- getURL("https://raw.githubusercontent.com/frankmoriarty/covid_ie/master/counties_population_cso_2016.csv")
cso <- read.csv(text = h, stringsAsFactors = FALSE)
  
n <- nrow(newcases)
the.date <- newcases$date[n]



```

## Counties most affected

As we can see in the map, **Dublin** and **Cavan** have suffered the highest relative proportion of cases, followed by Monaghan, Westmeath, Kildare and Wicklow. As of `r the.date` , it is mainly counties of the west that have been least affected



```{r, include=TRUE, message=FALSE}

library(sf)
library(tidyverse)
library(raster)
library(dplyr)
library(spData)
library(tmap)    # for static and interactive maps





library(csodata)

shp <- cso_get_geo("ac")




x <- getURL("https://raw.githubusercontent.com/frankmoriarty/covid_ie/master/doh_covid_ie_cases_analysis.csv")
newcases <- read.csv(text = x, stringsAsFactors=FALSE)


n <- nrow(newcases)

newcases$date<- as.factor(newcases$date)
newcases$date <- as.Date(newcases$date, format= "%d-%b")
the.date <- newcases$date[n]


counties_start <- grep("cases_carlow", colnames(newcases))
counties_stop <- grep("cases_wicklow", colnames(newcases))

#kept last 14 days
counties <- tibble(newcases[which(newcases$date >= the.date - 14 ), counties_start:counties_stop])
counties1 <- as.data.frame(counties)
N<- nrow(counties1)
#maybe no need to unlist : as.numeric(unlist(counties1))
counties1<- sapply(counties1,as.numeric)
counties2 <- counties1[N,] - counties1[N-13,]

counties3 <- data.frame(t(counties2[]))
#unused below
#colnames(counties3) <- counties2[, 1]


counties3 <- data.frame(r1=names(counties3), t(counties3))
colnames(counties3)[2] <- "cases"
colnames(counties3)[1] <- "county_doh"
#counties3$cases <- as.numeric(as.character(counties3$cases))
counties3[,1] <- sub("cases_", "", counties3[,1])
#counties3$cases <- as.numeric(counties3$cases)

county_cases <- merge(cso, counties3, by="county_doh")

county_cases$cases_per_mil <- 1000000*county_cases$cases/county_cases$pop

names(shp)[8] <- "county"
shp$county[20]="Dun Laoghaire-Rathdown"
county_cases$county[8]="Dun Laoghaire-Rathdown"

merged_cases <- merge(shp, county_cases, by="county")

tm_shape(merged_cases) + tm_fill(col="cases_per_mil", style = "pretty", title=c(paste("Cases per million", the.date))) + tm_borders() + tm_layout(asp = 1.28)

                                 

```   
                

