---
title: "Age-Groups Most Affected"
author: "Frank Moriarty and Ciaran Prendergast"
date: "5/11/2020"
output: html_document
---


```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = FALSE)

data1 <- read.csv("https://opendata-geohive.hub.arcgis.com/datasets/d8eb52d56273413b84b0187a4e9117be_0.csv?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D")

n <- nrow(data1)

data1$Date <- as.character(data1$Date)
data1$Date <- substr(data1$Date, 1, 10)
data1$Date <- as.Date(data1$Date, "%Y/%m/%d")
the.date <- data1$Date[n]
```


## Breakdown of Ages Affected

As we can see in the graph below, the most affected group has been the **Over 65s**, while people aged in their early 20s or younger remain the least likely to have contracted the virus. The disproportionatly fewer young people being affected scews the mean somewhat; as we can see when it is adjusted to include only people 15 years old or older. Updated as of `r the.date`.



```{r message=FALSE, warning=FALSE, , include=TRUE}
rm(list=ls())

library(tidyverse)
library(ggplot2)
data1 <- read.csv("https://opendata-geohive.hub.arcgis.com/datasets/d8eb52d56273413b84b0187a4e9117be_0.csv?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D")

n <- nrow(data1)
the.date <- data1$Date[n]
age_population_cso <- read.csv("https://raw.githubusercontent.com/frankmoriarty/covid_ie/master/age_population_cso_2019.csv")


totalllcases <- as.numeric(data1$TotalConfirmedCovidCases)
age <- c("0-4","05-14","15-24","25-34","35-44","45-54","55-64","65+")


cases <- as.numeric(c(data1$Aged1to4[which(data1$Date == the.date )],
                      data1$Aged5to14[which(data1$Date == the.date )],
                      data1$Aged15to24[which(data1$Date == the.date )],
                      data1$Aged25to34[which(data1$Date == the.date )],
                      data1$Aged35to44[which(data1$Date == the.date )],
                      data1$Aged45to54[which(data1$Date == the.date )],
                      data1$Aged55to64[which(data1$Date == the.date )],
                      data1$Aged65up[which(data1$Date == the.date )]))

#the cases by age all as numeric


popcso<- age_population_cso$pop_estimate[c(1:7,11)]      #the population estimates  


popcso.dividedbymillion <- popcso/1000000          #the population estimates divided by million as per formula in notebook



cases1M <- cases/popcso.dividedbymillion         #this gives a string of the cases per one million by age

df <- data.frame(age,cases1M)

the.plot <- ggplot(data=df, aes(x= age, y=cases1M,fill="indianred4", color="indianred4")) + geom_bar(stat='identity') 




total.cases <-totalllcases[which(data1$Date == the.date )]  #this gives the total cases on the date

totalpop.dividedby1m <- (age_population_cso$pop_estimate[which(age_population_cso$age_group=='cases_all_ages')])/1000000

mean.cases1M<- as.numeric(total.cases)/as.numeric(totalpop.dividedby1m)   #this is population mean for average abline


total.cases15plus <- totalllcases[which(data1$Date == the.date )] -  (( data1$Aged1to4[which(data1$Date == the.date )]+               data1$Aged5to14[which(data1$Date == the.date )])) #this gave the total amount of cases 15 plus

pop15plus.dividedby1M <- (age_population_cso$pop_estimate[which(age_population_cso$age_group=='cases_age_15_plus')])/1000000    #this was the population 15 plus divided by 1 million as per formula in notebook

mean.cases15plus1M <- total.cases15plus/pop15plus.dividedby1M  #this shows the total cases 15 plus per 1 million people



plot2 <- the.plot + scale_color_manual(values="indianred4",guide=FALSE)+scale_fill_manual(values="indianred4", guide=FALSE)






plot3 <- plot2 +geom_hline(aes(yintercept= mean.cases15plus1M, linetype = "Mean 15-plus"), colour= "green", size=1.25) +
  geom_hline(aes(yintercept=mean.cases1M , linetype = "Mean Total Population"), colour= "yellow", size=1.25) +
  scale_linetype_manual(name = "", values = c(2, 2), 
                        guide = guide_legend(override.aes = list(color = c("green", "yellow"))))
plot4 <- plot3 + labs(title= "Confirmed Covid Cases by Age",y= "Cases per Million", x = "Age-groups (Yr)")

plot4 +theme(legend.key = element_rect(fill = "lightblue", color = NA), axis.text=element_text(size=8, face = "bold"),
             axis.title=element_text(size=14,face="bold"), plot.title = element_text(size=16, face = "bold"))


```






