---
title: "Covid-Progression-over-Time"
author: "Frank Moriarty and Ciaran Prendergast"
date: "5/18/2020"
output: html_document
---


```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo=FALSE)
library(tidyverse)
library(RCurl)
x <- getURL("https://raw.githubusercontent.com/frankmoriarty/covid_ie/master/doh_covid_ie_cases_analysis.csv")
data1 <- read.csv(text = x)
n <- nrow(data1)
the.date <- data1$date[n]
```
 

## Progression of the Virus

As we can see here, the number of diagnosed Covid-19 cases in Ireland has grown exponentially, reaching **`r data1$cases_cumul[which(data1$date == the.date)]`** on `r the.date`. On the brightside, however, we can see that the rate of infection has fallen drastically; as the impact of *social distancing* translates into each infected person infecting fewer and fewer other people.


```{r Covid Progression, include=TRUE, message=FALSE, warning=FALSE}

library(tidyverse)
library(RCurl)
x <- getURL("https://raw.githubusercontent.com/frankmoriarty/covid_ie/master/doh_covid_ie_cases_analysis.csv")
data1 <- read.csv(text = x)



n <- nrow(data1)

total_cases <- as.numeric(data1$cases_cumul)

data1$newcases[1] <- NA
if (n > 1) for(i in 2:n) data1$newcases[i] <- ( total_cases[i] - total_cases[i-1] )


data1$newcases_per[1] <- NA
if (n > 1) for(i in 2:n) data1$newcases_per[i] <- ( 100*(total_cases[i] - total_cases[i-1])/total_cases[i-1] )





total1 <- c(1,1,1,2,6,13,18,19,21,24,34,43,70,90,129,169,data1$cases_cumul)  #this gives the total cases with the manually-typed numbers describing (constant) days not included in csv file

increase2 <- c(NA,NA,NA,NA,NA,NA,NA,5.6,10.5,14.3,41.7,26.5,62.8,28.6,43.3,31,60.36,na.omit(data1$newcases_per))  #this gives us the vector for increase on total, the nas are included to ensure that that the line only starts on the 7th
earlier.dates <- c("29-Feb", "1-Mar", "2-Mar", "3-Mar", "4-Mar", "5-Mar", "6-Mar", "7-Mar", "8-Mar", "9-Mar","10-Mar", "11-Mar", "12-Mar","13-Mar", "14-Mar", "15-Mar")
dates<- c(earlier.dates,data1$date)
dates0 <- as.factor(dates)                   #this calls the dates
dates1 <- as.Date(dates0, format= "%d-%b") 
df <- data.frame(increase2,dates1,total1)

ylim.prim <- c(0, 60)   
ylim.sec <- c(0, 18000)  

b <- diff(ylim.prim)/diff(ylim.sec)
a <- b*(ylim.prim[1] - ylim.sec[1])

plot <- ggplot(df, aes(dates1)) +
  geom_col(aes(y = increase2, color="tomoto4")) +
  geom_line(aes(y = a + total1*b, color="darkblue")) +
  scale_y_continuous("% Increase on Total", sec.axis = sec_axis(~ (. - a)/b, name = "Total Cases")) +
  scale_x_date("Dates") +
  ggtitle("Progression of Covid-19 in Ireland") 



plot2 <- plot + scale_color_discrete(name = "", labels = c("Total Cases", "% Increase on Total") )


plot2 + theme(legend.key = element_rect(fill = "lightblue", color = NA), axis.text=element_text(size=12, face = "bold"),
              axis.title=element_text(size=14,face="bold"), plot.title = element_text(size=16, face = "bold"))

```


